---
title: "Documentation on RHoMIS dataset"
author: "Romain Frelat"
date: "27-05-2024"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("Param_RHOMIS.R")
conv_F <- c(0.5, 0.75, 0.75, 0.86, 0.6) #female
conv_M <- c(0.5, 0.75, 0.925, 1, 0.73) #male
hhdb_rhomis <- readRDS("../Data/Processed/HHDB_RHOMIS_27052024.rds")
```

## Objectives

This document records the steps and choices made to transform the RHoMIS dataset into the format of the [farmhousehold Data Platform](https://github.com/rfrelat/farmhousehold). It goes together and complements the script file `Get_RHOMIS.R`. The original dataset is an updated version of:   
van Wijk, M., Hammond, J., Gorman, L., Adams, S., Ayantunde, A., Baines, D., ... & Yameogo, V. (2020). The Rural Household Multiple Indicator Survey, data from 13,310 farm households in 21 countries. Scientific Data, 7(1), 46.


## Crop table

The crop table contains 8 columns with information on crop cultivation and usage.

|  Name |         Definition        |   Unit   | 
|:------|:--------------------------|:---------|
| hhid  | household id              |          |
| name  | name of the crop          |          |
| land_area_ha| land cultivated     | hectare  |
| harvest_kg | amount harvested     | kg       |
| consumed_kg | amount consumed     | kg       |
| sold_kg | amount sold             | kg       |
| income_lcu | income from sells    | lcu      |


### Land area
The land cultivated area was transformed into hectare with the [rhomis package](https://github.com/RHoMIS/rhomis-R-package).  

We estimated the land area per crop from the reported proportion of land cultivated per crop (6 categories: all=90%, most=70%, half=50%, underhalf=20%, little=10%, none=0%). We corrected these percentages when necessary to make sure that the sum of the crop areas was equal to the area of land cultivated (sum of the proportions=100%). 

For households without proportions per crop, we assumed that the proportion of crop area is equal to the proportion of quantities harvested.


### Crop diversity
We recorded all crops cultivated by farmers. When crops were only listed as grown, without any information on production or usage (not considered as a main crop), we only recorded the hhid and the name in the table, all other columns were filled with NA.


\newpage
## Livestock table

The livestock table contains 3 columns with information on livestock herd

|  Name |         Definition        |   Unit   | 
|:------|:--------------------------|:---------|
| hhid  | household id              |          |
| name  | name of the livestock     |          |
| n     | number of livestock kept  |          |


The livestock are in 16 categories:
bees, buffalo, camel, cattle, chicken, dogs, donkeys_horses, duck, fish, goats, guinea_pigs, otherpoultry, oxen, pigs, rabbits, sheep.   

  
## Livestock production table
The livestock production table contains 7 columns with information on livestock productions:  

|  Name |         Definition        |   Unit   | 
|:------|:--------------------------|:---------|
| hhid  | household id              |          |
| name  | name of the livestock     |          |
| prod  | livestock production      |          |
| harvest_kg | amount harvested     | kg       |
| consumed_kg | amount consumed     | kg       |
| sold_kg | amount sold             | kg       |
| income_lcu | income from sells    | lcu      |


We consider the production of meat, milk, eggs and honey as well as whole animal sales.


### Whole animal sales
For the sales of whole animals, the quantities were given in number of animals. We converted the numbers into kg using the TLU conversion factors (1TLU=250kg). As the animal were sold alive, no quantity was consumed, the quantity harvested equals the quantity sold. 

```{r conv_tlu}
conv_tlu[conv_tlu>0]*250
```

\newpage
## Household information table
The initial household information table contains information on household composition, off farm activities, and food security.


|  Name  |         Definition       | Unit   | 
|:-------|:-------------------------|:-------|
| hhid  | household id              |          |
| country  | country of the survey      |          |
| year  | year of the survey      |          |
| gps_lat | latitude in decimal degrees | °N       |
| gps_lon | longitude in decimal degrees | °E       |
| hh_size_members | size of the household in number of persons |     |
| hh_size_mae | size of the household in male adult equivalent | MAE |
| head_age | age of the household head |     |
| head_gender | gender of the household head | 'f' or 'm' |
| off_farm_lcu | off farm income per year    | lcu      |
| off_farm_div | diversity of off farm activities   |       |
| hdds | household diet diversity score based on 10 groups    |       |
| fies | Food Insecurity Experience Scale based on 8 questions   |       |
| foodshortage_count | number of months with food shortage    |       |
| foodshortage_months | name of the months with food shortage    |       |
| currency_conversion | conversion from local currency to power parity purchase usd    | lcu/usd        |

### General information
`hhid`is made of the variable `id_unique`. This variable is a unique identifier for the household that can connect all tables (crop, livestock, and livestock production).  

The GPS coordinates are provided with a rounding at 0.1 decimal degree. The currency power parity purchase conversion factor is provided by the World Bank depending on the country and the year of the survey.  

### Household size
The household composition is converted by the [rhomis package](https://github.com/RHoMIS/rhomis-R-package). The male adult equivalent are calculated with 5 age-groups: 0-4, 5-10, 11-24, 25-50, 51+ and with the corresponding coefficient per gender and age class. 


### Off farm income
The household composition is estimated from the [rhomis package](https://github.com/RHoMIS/rhomis-R-package) based on the farm production and the estimated percentage of income from off-farm activities.

### Food security
We used the food security indicators processed by the [rhomis package](https://github.com/RHoMIS/rhomis-R-package).


\newpage
## Crop and Livestock summary

The household information table also contains 28 columns with summary information from crop and livestock tables. 

|  Name   |         Definition       |  Unit| 
|:--------|:-------------------------|:-----|
| hhid  | household id              |          |
| land_cultivated_ha  | total land cultivated | ha          |
| crop_div  | number of crop cultivated      |          |
| crop_name | names of crop cultivated       |        |
| crop_harvest_kg | total crop harvest       | kg       |
| crop_yield_kg_per_ha | crop yield          | kg/ha     |
| crop_sold_kg | quantity of crop sold       | kg |
| crop_sold_perc | percentage of quantities of crop sold    | %      |
| crop_income_div | number of different crop sold   |       |
| crop_income_lcu | total income from crop production  | lcu   |
| crop_value_lcu | value of crop produced but not sold | lcu   |
| crop_consumed_kcal | energy value from crop consumed   | kcal  |
| livestock_tlu | herd size              | tlu      |
| lstk_div | number of livestock species herded     |        |
| lstk_name | names of livestock species herded |      |
| lstk_harvest_kg | total livestock product harvested     | kg |
| lstk_sold_kg | quantity of livestock product sold       | kg |
| lstk_sold_perc | percentage of livestock production sold    | %      |
| lstk_income_div | number of different livestock products sold   |       |
| lstk_income_lcu | total income from livestock production    | lcu      |
| lstk_value_lcu |  value of livestock production not sold | lcu   |
| lstk_consumed_kcal | energy value from livestock consumed   | kcal  |
| farm_div | number of crop and livestock species     |        |
| farm_harvest_kg | total farm production     | kg |
| farm_sold_perc_kg | percentage of farm production sold    | %      |
| farm_income_div | number of different farm products sold   |       |
| farm_income_lcu | total income from farm production    | lcu      |
| tot_income_lcu |  total income (farm + off farm) | lcu   |
| farm_consumed_kcal | energy value from farm production consumed   | kcal  |
| off_farm_perc | percentgae of income from off farm activities   | %      |

All the values are calculated automatically with the function `calc_farm_prod()`. The calculations are simple and summarize, per household, the quantities reported in the crop and livestock tables.  

For energy conversion, we used estimates mostly from the FoodData Central of the U.S.Departement of Agriculture (https://fdc.nal.usda.gov/). Below are the energy conversion factors for the twenty most important crops:

```{r conv_kcal}
top20 <- sort(tapply(hhdb_rhomis$crop$harvest_kg, hhdb_rhomis$crop$name, sum, na.rm=TRUE), decreasing = TRUE)[1:20]
top_conv_energy <- hhdb_rhomis$conv_energy[sort(names(top20))]
top_conv_energy[top_conv_energy>0 & !is.na(top_conv_energy)]
```

\newpage
## GIS information

Based on the GPS coordinates of households, we extracted:  

- the Dixon farming system classification for Sub-Saharan Africa [Dixon et al. 2021](https://www.routledge.com/Farming-Systems-and-Food-Security-in-Africa-Priorities-for-Science-and-Policy-Under-Global-Change/Dixon-Garrity-Boffa-Williams-Amede-Auricht-Lott-Mburathi/p/book/9781032082141)  
- the population density estimated by the [Gridded Population of the World (GPWv4)](https://sedac.ciesin.columbia.edu/data/collection/gpw-v4/documentation) and download from [geodata package](https://github.com/rspatial/geodata)   
- the travel time to cities estimated by [Nelson et al. 2019](https://www.nature.com/articles/s41597-019-0265-5) and download from [geodata package](https://github.com/rspatial/geodata)   
- the Koeppen's Climate Classification from:  
Beck, H.E., et al. (2018) "Present and future Köppen-Geiger climate classification maps at 1-km resolution", *Nature Scientific Data*, 5, 180214 [DOI 10.1038/sdata.2018.214](https://doi.org/10.1038/sdata.2018.214)  


## Summary

We kept only *rural* households with at least one crop or one livestock. In total, the dataset is made of `r nrow(hhdb_rhomis$hhinfo)` households, with information on `r nrow(hhdb_rhomis$crop)` crop cultivated and `r nrow(hhdb_rhomis$lstk)` livestock species herd. The dataset made of the four tables, together with tlu and energy conversion factors are binded together into a `farmhousehold` object and saved into the file `HHDB_RHOMIS_27052024.rds`.
