---
title: "Documentation SIMLESA 2013"
author: "Romain Frelat"
date: "27-05-2024"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("Param_SIMLESA_MWI2013.R")
hhdb_simlesa_mwi2013 <- readRDS("../Data/Processed/HHDB_SIMLESA_MWI2013.rds")
```

## Objectives

This document records the steps and choices made to transform SIMLESA Malawi 2013 dataset into the format of the Household Data Platform. It goes together and complements the script file `Get_SIMLESA_MWI2013.R`. The original dataset is:   

Pathways to sustainable intensification in Eastern and Southern Africa - Malawi 2013
Marenya, Paswel; Kassie, Menale; Muricho, Geoffrey; Alemu, Solomon; Yirga, Chilot; Mishili, Fulgence; Obare, Gideon; Tostao, Emilio, 2016, "Pathways to sustainable intensification in Eastern and Southern Africa - Malawi 2013", https://hdl.handle.net/11529/10760, CIMMYT Research Data & Software Repository Network, V3, UNF:5:ascjPZI1bFZMaMFDIMe7uA


## Crop table

The crop table contains 7 columns with information on crop cultivation and usage.

|  Name |         Definition        |   Unit   | 
|:------|:--------------------------|:---------|
| hhid  | household id              |          |
| name  | name of the crop          |          |
| land_area_ha| land cultivated     | hectare  |
| harvest_kg | amount harvested     | kg       |
| consumed_kg | amount consumed     | kg       |
| sold_kg | amount sold             | kg       |
| income_lcu | income from sells    | lcu      |


The land and crop management are described in Module 4 Part A, the utilization of crop harvested in Module 4 Part B, and the incomes in Module 4 Part C.  


### Land area   

We estimated the land area per crop from the questions in module 4, part A.

The field size were already provided in hectare. We used the reported percentage of the sub-plot cover by the three crops to estimate the area per crop. We corrected these percentages when necessary to make sure that the sum of the crop areas per sub-plot is not higher than the sub-plot area. When the percentage was not reported, we estimated it based on the ratio of yields among the three main crops of the sub-plot. The remaining area (if any) was attributed to tree crop planted on the same field. Yet in most cases the tree crop (mostly banana and coffee) don´t have proper land area (nor information about harvested quantities).  


### Harvest and uses   

The quantities of crop harvested are provided both in module 4 part A and part B. We decided to use the quantity of crop harvested provided in part B in priority (unless it was null, in this case we used the quantity reported in part A), because it was easily related to crop uses. Similarly for the quantity of crop sold is provided in module 4 part B and C and we decided to use the quantity provided in part B in priority.

At the end, the quantities were checked for consistency. When the amount consumed and sold was higher than the amount harvested, we used the sum of the amount consumed and sold as quantity harvested, unless the resulting yield was higher than 20t/ha. In such case (abnormally high quantities sold and consumed), we lower the quantities sold and consumed using the same proportions as originally reported but calculated on the amount harvested.


## Livestock table

The livestock table contains 3 columns with information on livestock herd

|  Name |         Definition        |   Unit   | 
|:------|:--------------------------|:---------|
| hhid  | household id              |          |
| name  | name of the livestock     |          |
| n     | number of livestock kept  |          |


The information on the herd structure is provided in Module 10 part E: Livestock ownership.   
We grouped *Crossbred/exotic*, *Heifers*, *Bulls*, *Calves*, *Oxen*, *Indigenous cows* into **cattle**.  
We grouped *Donkeys*, *Horses*, and *Mule* into **Donkeys_horses**.  



## Livestock production table

The livestock production table contains 7 columns with information on livestock productions:  

|  Name |         Definition        |   Unit   | 
|:------|:--------------------------|:---------|
| hhid  | household id              |          |
| name  | name of the livestock     |          |
| prod  | livestock production      |          |
| harvest_kg | amount harvested     | kg       |
| consumed_kg | amount consumed     | kg       |
| sold_kg | amount sold             | kg       |
| income_lcu | income from sells    | lcu      |

The information on milk, dairy products, eggs and honey are captured in Module 10 part E: Livestock ownership.  

The quantities consumed or harvest was not reported. So we used the quantities sold as quantity harvested and the quantities consumed was set to 0.  

The income was calculated as the product of the quantity sold (A6) and the price per kg (A7). We checked these values by calculating the median price per item per kg, and we used the reported price, unless it was too different than the median price (=5 times higher or lower) and the value reported in `valsold` was closer to the median price.  

The quantities sold (in numbers) are transformed in kg using the conversion of one egg to `r conv_egg` kg, and the livestock sold alive using the TLU conversion factors (1TLU=250kg).

```{r conv_tlu}
hhdb_simlesa_mwi2013$conv_tlu*250
```
  



## Household information table
The household information table contains information on household composition, off farm activities, and food security.


|  Name  |         Definition       | Unit   | 
|:-------|:-------------------------|:-------|
| hhid  | household id              |          |
| country  | country of the survey      |          |
| year  | year of the survey      |          |
| gps_lat | latitude in decimal degrees | °N       |
| gps_lon | longitude in decimal degrees | °E       |
| hh_size_members | size of the household in number of persons |     |
| hh_size_mae | size of the household in male adult equivalent | MAE |
| head_age | age of the household head |     |
| head_gender | gender of the household head | 'f' or 'm' |
| off_farm_lcu | off farm income per year    | lcu      |
| off_farm_div | diversity of off farm activities   |       |
| hdds | household diet diversity score based on 10 groups    |       |
| fies | Food Insecurity Experience Scale based on 8 questions   |       |
| foodshortage_count | number of months with food shortage    |       |
| foodshortage_months | name of the months with food shortage    |       |
| currency_conversion | conversion from local currency to power parity purchase usd    | lcu/usd        |


### General information   

`hhid`is made of the prefix "SIMLESA_MWI_2013" and the variable `hhid`. This variable is a unique identifier for the household that can connect all the tables (crop, livestock, and livestock production).  

The GPS coordinates were not provided. Therefore, we used the centroid of the reported district or EPA as coordinates of the households. 

The currency power parity purchase conversion factor is provided by the World Bank for Malawi in 2013 (= `r round(conv_lcu,2)`mwk/usd).  

### Household size    

The household composition is provided in Module 2 part A. The male adult equivalent are calculated with 5 age-groups: 0-4, 5-10, 11-24, 25-50, 51+ and with the corresponding coefficient per gender and age class. We used the reported age in month when the age in years was missing (for babies).

```{r conv_mae}
conv_mae <- rbind("Male"=conv_M, "Female"=conv_F)
colnames(conv_mae) <- c("0-4", "5-10", "11-24", "25-50", "51+")
conv_mae
```

### Off farm income    

We considered the off farm income reported in Module 11: Other sources of household income.  

\newpage
### Food security   

The diet diversity was calculated from the 7-day recall of food consumption (module 8). We grouped the 18 groups into 10 groups (to be similar to the LSMS-ISA Malawi survey).
```{r cat_hdds}
cbind("Original"=substr(names(cat_hdds),1,50), "Merged"=gsub("HDDS_", "", as.character(cat_hdds)))
```

\newpage
## Crop and Livestock summary

The household information table also contains 28 columns with summary information from crop and livestock tables. 

|  Name   |         Definition       |  Unit| 
|:--------|:-------------------------|:-----|
| hhid  | household id              |          |
| land_cultivated_ha  | total land cultivated | ha          |
| crop_div  | number of crop cultivated      |          |
| crop_name | names of crop cultivated       |        |
| crop_harvest_kg | total crop harvest       | kg       |
| crop_yield_kg_per_ha | crop yield          | kg/ha     |
| crop_sold_kg | quantity of crop sold       | kg |
| crop_sold_perc | percentage of quantities of crop sold    | %      |
| crop_income_div | number of different crop sold   |       |
| crop_income_lcu | total income from crop production  | lcu   |
| crop_value_lcu | value of crop produced but not sold | lcu   |
| crop_consumed_kcal | energy value from crop consumed   | kcal  |
| livestock_tlu | herd size              | tlu      |
| lstk_div | number of livestock species herded     |        |
| lstk_name | names of livestock species herded |      |
| lstk_harvest_kg | total livestock product harvested     | kg |
| lstk_sold_kg | quantity of livestock product sold       | kg |
| lstk_sold_perc | percentage of livestock production sold    | %      |
| lstk_income_div | number of different livestock products sold   |       |
| lstk_income_lcu | total income from livestock production    | lcu      |
| lstk_value_lcu |  value of livestock production not sold | lcu   |
| lstk_consumed_kcal | energy value from livestock consumed   | kcal  |
| farm_div | number of crop and livestock species     |        |
| farm_harvest_kg | total farm production     | kg |
| farm_sold_perc_kg | percentage of farm production sold    | %      |
| farm_income_div | number of different farm products sold   |       |
| farm_income_lcu | total income from farm production    | lcu      |
| tot_income_lcu |  total income (farm + off farm) | lcu   |
| farm_consumed_kcal | energy value from farm production consumed   | kcal  |


All the values are calculated automatically with the function `calc_farm_prod()`. The calculations are simple and summarize per households the quantities reported in the crop and livestock tables.  


For energy conversion, we used estimates mostly from the FoodData Central of the U.S.Departement of Agriculture (https://fdc.nal.usda.gov/). Below are the energy conversion factors:
```{r conv_kcal}
top20 <- sort(tapply(hhdb_simlesa_mwi2013$crop$harvest_kg, hhdb_simlesa_mwi2013$crop$name, sum, na.rm=TRUE), decreasing = TRUE)[1:20]
top_conv_energy <- hhdb_simlesa_mwi2013$conv_energy[sort(names(top20))]
top_conv_energy[top_conv_energy>0 & !is.na(top_conv_energy)]
```

\newpage
## GIS information   

Based on the GPS coordinates of centroid of the district (or EPA) of the households, we extracted:  
- the Dixon farming system classification for Sub-Saharan Africa [Dixon et al. 2021](https://www.routledge.com/Farming-Systems-and-Food-Security-in-Africa-Priorities-for-Science-and-Policy-Under-Global-Change/Dixon-Garrity-Boffa-Williams-Amede-Auricht-Lott-Mburathi/p/book/9781032082141)
- the Koeppen's Climate Classification from:  
Beck, H.E., et al. (2018) "Present and future Köppen-Geiger climate classification maps at 1-km resolution", *Nature Scientific Data*, 5, 180214 [DOI 10.1038/sdata.2018.214](https://doi.org/10.1038/sdata.2018.214) 
The population density, and the travel time to cities could not be estimated based on these aggregated coordinates.


## Summary

We kept only *rural* households with at least one crop or one livestock. In total, the dataset is made of `r nrow(hhdb_simlesa_mwi2013$hhinfo)` households, with information on `r nrow(hhdb_simlesa_mwi2013$crop)` crop cultivated and `r nrow(hhdb_simlesa_mwi2013$lstk)` livestock species. The dataset made of the four tables, together with tlu and energy conversion factors are binded together into a `farmhousehold` object and saved into the file `HHDB_SIMLESA_MWI2013.rds`.  

