---
title: "Documentation LSMS-ISA Malawi 2019"
author: "Romain Frelat"
date: "27-05-2024"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("Param_LSMS_MWI2019.R")
hhdb_lsms_mwi2019 <- readRDS("../Data/Processed/HHDB_LSMS_MWI2019.rds")
```

## Objectives

This document records the steps and choices made to transform LSMS-ISA Malawi 2019 dataset into the format of the Household Data Platform. It goes together and complements the script file `Get_LSMS_MWI2019.R`. The original dataset (MWI_2019_IHS-V_v05_M, DOI 10.48529/yqn3-zv74) can be accessed from the World Bank microdata library:
https://microdata.worldbank.org/index.php/catalog/3818/ [last accessed 30/11/2023]


## Crop table

The crop table contains 8 columns with information on crop cultivation and usage.

|  Name |         Definition        |   Unit   | 
|:------|:--------------------------|:---------|
| hhid  | household id              |          |
| name  | name of the crop          |          |
| land_area_ha| land cultivated     | hectare  |
| harvest_kg | amount harvested     | kg       |
| consumed_kg | amount consumed     | kg       |
| sold_kg | amount sold             | kg       |
| income_lcu | income from sells    | lcu      |
| use_list | list all usages        | categories |


The crop information is divided in three different sections: the rainy season (module C and G), the dry season (module K and M), and the perenial crops (module P). The same transformations were made for the three sources of information, that were merged together after processing.

### Land area  

The field size were transformed into hectare. Most were measured by GPS and reported in $acre$ ($1acre=0.404686ha$). When not available, we used the household estimates and the following conversion factors.
```{r conv_ha}
conv_ha
```

Additionally, we used the reported percentage of the field cover by each crop (given in 5 categories) to estimate the area per crop. We corrected these percentages when necessary to make sure that the sum of the crop areas per field is not higher than the field area. 


### Harvest and uses  

The crop harvest and usage are estimated from the post harvest questionnaire, in section 11. The conversion factors for units of weight are calculated from the file `ihs_seasonalcropconversion_factor_2020.csv`, `ihs_foodconversion_factor_2020.csv`, and `ihs_treeconversion_factor_2020.csv`. We simplified and used only one conversion factor per unit (the median), without considering the different crop. 

```{r conv_kg, eval=FALSE}
#the list is much too big
conv_kg
```

The amount of crop consumed was not reported in the survey for the dry season and the perenial crop. In these caases, we calculated it as the difference between the quantity harvested and the reported quantity used (sum of the quantities given as gift, as payment of debts, to feed livestock, for by-product fabrication, for seed production and lost).  

The quantities are checked for consistency. When the amount consumed and sold was higher than the amount harvested, we used the sum of the amount consumed and sold as quantity harvested, unless the resulting yield was higher than 20t/ha. In such case (abnormally high quantities sold and consumed), we lower the quantities sold and consumed using the same proportions as originally reported but calculated on the amount harvested.

## Livestock table

The livestock table contains 3 columns with information on livestock herd

|  Name |         Definition        |   Unit   | 
|:------|:--------------------------|:---------|
| hhid  | household id              |          |
| name  | name of the livestock     |          |
| n     | number of livestock kept  |          |


The information on the herd structure is provided in section 8 of livestock questionnaire. 

We grouped *local-hen*, *local-cock*, *chicken-layer/chicken-broiler* and *chicks* into **poultry**.  
We grouped *cow*, *bull*, *ox*, *calf*, *steer/heifer*, and *improved* into **cattle**.  
We grouped *hare*, *hare and mbira*, and *mbila* into **rabbit**.  

\newpage
## Livestock production table
The livestock production table contains 7 columns with information on livestock productions:  

|  Name |         Definition        |   Unit   | 
|:------|:--------------------------|:---------|
| hhid  | household id              |          |
| name  | name of the livestock     |          |
| prod  | livestock production      |          |
| harvest_kg | amount harvested     | kg       |
| consumed_kg | amount consumed     | kg       |
| sold_kg | amount sold             | kg       |
| income_lcu | income from sells    | lcu      |


### Milk, egg and manure  

The livestock production is recorded in module S, recording chicken eggs, guinea fowl eggs, cow milk, manure, meat, and skins and hides production. For meat, manure, and skin and hides, we attributed the production to the livestock category with the largest herd (converted in TLU number).


```{r conv_prod}
conv_lstk_prod
```



### Whole animals  

The selling of whole animals (reported in module R1) is also considered as livestock product. The whole animals are always sold (no quantities consumed, the quantity harvested equals the quantity sold). We converted them into kg using the TLU conversion factors (1TLU=250kg).

```{r conv_tlu}
conv_tlu*250
```

\newpage
## Household information table
The initial household information table contains information on household composition, off farm activities, and food security.


|  Name  |         Definition       | Unit   | 
|:-------|:-------------------------|:-------|
| hhid  | household id              |          |
| country  | country of the survey      |          |
| year  | year of the survey      |          |
| gps_lat | latitude in decimal degrees | °N       |
| gps_lon | longitude in decimal degrees | °E       |
| hh_size_members | size of the household in number of persons |     |
| hh_size_mae | size of the household in male adult equivalent | MAE |
| head_age | age of the household head |     |
| head_gender | gender of the household head | 'f' or 'm' |
| off_farm_lcu | off farm income per year    | lcu      |
| off_farm_div | diversity of off farm activities   |       |
| hdds_score | household diet diversity score based on 10 groups    |       |
| foodshortage_count | number of months with food shortage    |       |
| foodshortage_months | name of the months with food shortage    |       |
| currency_conversion | conversion from local currency to power parity purchase usd    | lcu/usd        |


### General information  

`hhid`is made of the prefix "LSMS_MWI_2019" and the variable `case_id`. This variable is a unique identifier for the household that can connect all tables (crop, livestock, and livestock production).  

The GPS coordinates are provided with a rounding at 0.1 decimal degree. The currency power parity purchase conversion factor is provided by the World Bank for Ethiopia in 2018 (= `r round(conv_lcu,2)`mwk/usd).  

### Household size  

The household composition is provided in section 1 of the household questionnaire. The male adult equivalent are calculated with 5 age-groups: 0-4, 5-10, 11-24, 25-50, 51+ and with the corresponding coefficient per gender and age class. 

```{r conv_mae}
conv_mae <- rbind("Male"=conv_M, "Female"=conv_F)
colnames(conv_mae) <- c("0-4", "5-10", "11-24", "25-50", "51+")
conv_mae
```

### Off farm income  

We considered the employment (reported in section 4 of the household questionnaire) or businesses (reported in section 12). For employment, we calculated the number of hours worked, and the salary per hour based on the last payment. If the salary was higher than $1000birr/h$ we set it to $1000birr/h$. Then we multiplied the number of hours worked by the salary per hour to get the annual salary. If the annual salary resulted to be lower than the reported payment, we reported the payment as annual salary.

### Food security  

The diet diversity was calculated from the 7-day recall of food consumption (module G).

The months with food shortage was directly retrieve from questions in module H.

## Crop and Livestock summary

The household information table also contains 28 columns with summary information from crop and livestock tables. 

|  Name   |         Definition       |  Unit| 
|:--------|:-------------------------|:-----|
| hhid  | household id              |          |
| land_cultivated_ha  | total land cultivated | ha          |
| crop_div  | number of crop cultivated      |          |
| crop_name | names of crop cultivated       |        |
| crop_harvest_kg | total crop harvest       | kg       |
| crop_yield_kg_per_ha | crop yield          | kg/ha     |
| crop_sold_kg | quantity of crop sold       | kg |
| crop_sold_perc | percentage of quantities of crop sold    | %      |
| crop_income_div | number of different crop sold   |       |
| crop_income_lcu | total income from crop production  | lcu   |
| crop_value_lcu | value of crop produced but not sold | lcu   |
| crop_consumed_kcal | energy value from crop consumed   | kcal  |
| livestock_tlu | herd size              | tlu      |
| lstk_div | number of livestock species herded     |        |
| lstk_name | names of livestock species herded |      |
| lstk_harvest_kg | total livestock product harvested     | kg |
| lstk_sold_kg | quantity of livestock product sold       | kg |
| lstk_sold_perc | percentage of livestock production sold    | %      |
| lstk_income_div | number of different livestock products sold   |       |
| lstk_income_lcu | total income from livestock production    | lcu      |
| lstk_value_lcu |  value of livestock production not sold | lcu   |
| lstk_consumed_kcal | energy value from livestock consumed   | kcal  |
| farm_div | number of crop and livestock species     |        |
| farm_harvest_kg | total farm production     | kg |
| farm_sold_perc_kg | percentage of farm production sold    | %      |
| farm_income_div | number of different farm products sold   |       |
| farm_income_lcu | total income from farm production    | lcu      |
| tot_income_lcu |  total income (farm + off farm) | lcu   |
| farm_consumed_kcal | energy value from farm production consumed   | kcal  |


All the values are calculated automatically with the function `calc_farm_prod()` from the package `farmhousehold`. The calculations are simple and summarize per households the quantities reported in the crop and livestock tables.  


For energy conversion, we used estimates mostly from the FoodData Central of the U.S.Departement of Agriculture (https://fdc.nal.usda.gov/). Below are the energy conversion factors:
```{r conv_kcal}
top20 <- sort(tapply(hhdb_lsms_mwi2019$crop$harvest_kg, hhdb_lsms_mwi2019$crop$name, sum, na.rm=TRUE), decreasing = TRUE)[1:20]
top_conv_energy <- hhdb_lsms_mwi2019$conv_energy[sort(names(top20))]
top_conv_energy[top_conv_energy>0 & !is.na(top_conv_energy)]
```

\newpage
## GIS information

Based on the GPS coordinates of households, we extracted:  

- the administrative region and district (from GADM)   
- the Dixon farming system classification for Sub-Saharan Africa [Dixon et al. 2021](https://www.routledge.com/Farming-Systems-and-Food-Security-in-Africa-Priorities-for-Science-and-Policy-Under-Global-Change/Dixon-Garrity-Boffa-Williams-Amede-Auricht-Lott-Mburathi/p/book/9781032082141)  
- the population density estimated by the [Gridded Population of the World (GPWv4)](https://sedac.ciesin.columbia.edu/data/collection/gpw-v4/documentation)    
- the travel time to cities estimated by [Nelson et al. 2019](https://www.nature.com/articles/s41597-019-0265-5)   
- the Koeppen's Climate Classification from:  
Beck, H.E., et al. (2018) "Present and future Köppen-Geiger climate classification maps at 1-km resolution", *Nature Scientific Data*, 5, 180214 [DOI 10.1038/sdata.2018.214](https://doi.org/10.1038/sdata.2018.214)   


## Summary

We kept only *rural* households with at least one crop or one livestock. In total, the dataset is made of `r nrow(hhdb_lsms_mwi2019$hhinfo)` households, with information on `r nrow(hhdb_lsms_mwi2019$crop)` crop cultivated and `r nrow(hhdb_lsms_mwi2019$lstk)` livestock species herd. The dataset made of the four tables, together with tlu and energy conversion factors are binded together into a `farmhousehold` object and saved into the file `HHDB_LSMS_MWI2019.rds`.  

